Monte Carlo calculation to characterize neutron and gamma fields for ANITA neutron facility at TSL

### 2015-04-17 Paper2015.tex
Condensed modelling section and updated early part of results section.

\section{Modelling}

A detailed model of the ANITA geometry was implemented in Geant4.
This is illustrated in overview in Fig.~\ref{fig:ANITAoverview}
Description of the geometry is given in \cite{Prokofiev2009,Prokofiev14}; a summary only is given here.
The spallation target is a tungsten cylinder of diameter \SI{5}{\cm} and length \SI{2.4}{\cm}.
The target is cooled by water and surrounded by a stainless-steel cooling jacket.
The target assembly with its cooling jacket appears as a small blue rectangle in Fig.~\ref{fig:ANITAoverview}.
Lead shielding blocks in the target region are also modelled, and shown in red Fig.~\ref{fig:ANITAoverview}.
A large electromagnet surrounding the target is modelled and shown in Fig.~\ref{fig:ANITAoverview} in grey (iron) and yellow (copper).
An iron collimator with variable apertures is modelled  and shown downstream of the target (grey).
For brevity in this summary we only present results for the standard collimator size (\SI{10.2}{\cm} diameter); in the final paper we will present results for cylindrical collimators in the range from \SIrange{3}{30}{\cm} diameter.

@@

Simulations were conducted using Geant4 version 10.0.
The binary intranuclear cascade model was selected in accordance with the results of earlier work~\cite{Platt13}, in which it performed well in simulations of neutron spallation in the region from \SIrange{0}{0}{\MeV}.\todo{Enter neutron energy range}
In the simulation protons were incident axially at an energy of \SI{180}{\MeV}.
Resulting gamma and neutron fields were evaluated at several locations.
In this paper results are presented for three of these: The Standard User Position (SUP), \SI{2.5}{\m} downstream of the target centre\todo{check centre}~\cite{Prokofiev2009}, the Close User Position (CUP), \SI{0.75}{\m} downstream of the target, and the CUP-TOF, \SI{0.84}{m} downstream of the target.\todo{Enter CUP \& CUP-TOF ranges}
The SUP, CUP and CUP-TOF positions are visible in blue in Fig.~\ref{fig:ANITAoverview}.
The SUP and CUP positions are of interest as they are positions used for SEE tests; the CUP-TOF position is the location of thin-film breakdown counter (TFBC) detectors used for beam monitoring and characterisation using time-of-flight (TOF) techniques.
Model validation will be demonstrated against CUP-TOF data (section~\ref{tbd}).

%------------------------------------------------
\section{Results}

In previous work Platt et al.~\cite{Platt13} reported simulations of a naked ANITA target, demonstrating good qualitative agreement with independent calculations and measurements of the fast neutron field at ANITA SUP~\cite{Prokofiev2009}.
Part of the motivation for this work was to investigate whether improvements in the fidelity of the geometry used in the Geant4 model would lead to increased quantitative agreement.\todo{Review first part of Results section -- move to introduction?}

\subsection{Neutron fluence rate, spatial distribution and spectrum}

Fig.~\ref{fig:SUPDensity} shows neutron spatial distribution at the SUP.
The collimation effect is clearly visible.
Fig.~\ref{fig:SUPProfile} compares simulated and measured neutron fluence profile at the SUP.
Measurements were taken with a TFBC in which \U238\ is the target material~\cite{Prokofiev2009}; accordingly, simulated results were folded in energy with the neutron fission cross section for \U238~\cite{tbd}.
The results show~\todo{Explain SUP profile results. Include explanation of ``umbra'' and ``penumbra''}
The predicted neutron fluence rate above \SI{10}{\MeV} at a proton current of \SI{200}{\nA} is \SI{0}{\neutron\per\cm\squared\per\second}, compared to \SI{0}{\neutron\per\cm\squared\per\second} from earlier calculations~\cite{Platt13} and \SI{0}{\neutron\per\cm\squared\per\second} from the parameterisation~\cite{Prokofiev2009}.
The parameterisation therefore exceeds the current calculations by tbd\%.\todo{Quantify integral yield}

\begin{figure}[t]
    \includegraphics[width=3in]{SUP10ColSpatialDistribution10MeVRADECS.png}
% \todo{Show density of neutrons above \SI{10}{\MeV} at SUP}
    \caption{Spatial distribution of neutrons above \SI{10}{\MeV} at the Standard User Position.}
    \label{fig:SUPDensity}
\end{figure}

\begin{figure}[t]
    \includegraphics[width=3in]{SUP10beamproFoldingRADECS.png}
% \todo{Show SUP profile, compared to measurements}
    \caption{Neutron profile at the Standard User Position}
    \label{fig:SUPProfile}
\end{figure}

>>>>

\section{Modelling}
Simulations were conducted using Geant4 version 10.0 and the binary intranuclear cascade model, as recommended in~\cite{Platt13}.
A detailed model of the ANITA geometry was implemented in Geant4, as illustrated in Fig.~\ref{fig:ANITAoverview}
Description of the geometry is given in \cite{Prokofiev2009,Prokofiev14}.
%; a summary only is given here.
%The spallation target is a tungsten cylinder of diameter \SI{5}{\cm} and length \SI{2.4}{\cm}.
%The target is cooled by water and surrounded by a stainless-steel cooling jacket.
%The target assembly with its cooling jacket appears as a small blue rectangle in Fig.~\ref{fig:ANITAoverview}.
%Lead shielding blocks in the target region are also modelled, and shown in red Fig.~\ref{fig:ANITAoverview}.
%A large electromagnet surrounding the target is modelled and shown in Fig.~\ref{fig:ANITAoverview} in grey (iron) and yellow (copper).
%An iron collimator with variable apertures is modelled  and shown downstream of the target (grey).
Except where indicated otherwise, all results in this summary are for the standard collimator size (\SI{10.2}{\cm} diameter).
In the final paper we will present results for cylindrical collimators in the range from \SIrange{3}{30}{\cm} diameter.
Simulated protons were incident axially at an energy of \SI{180}{\MeV}.
Resulting gamma and neutron fields were evaluated at several detector locations.
In this paper results are presented for three of these: the Standard User Position (SUP), \SI{2.5}{\m} downstream of the target centre\todo{check centre}~\cite{Prokofiev2009}, the Close User Position (CUP), \SI{0.75}{\m} downstream of the target, and the CUP-TOF, \SI{0.84}{m} downstream of the target.
The SUP and CUP positions are of interest as they are positions used for SEE tests; the CUP-TOF position is the location of thin-film breakdown counter (TFBC) detectors used for beam monitoring and characterisation using time-of-flight (TOF) techniques.
Model validation will be demonstrated against CUP-TOF data (section~\ref{tbd}).

@@

\section{Results}
In previous work Platt et al.~\cite{Platt13} reported simulations of a naked ANITA target, demonstrating good qualitative agreement with independent calculations and measurements of the fast neutron field at ANITA SUP~\cite{Prokofiev2009}.
Part of the motivation for this work was to investigate whether improvements in the fidelity of the geometry used in the Geant4 model would lead to increased quantitative agreement.\todo{Review first part of Results section -- move to introduction?}

\subsection{Neutron fluence rate, spatial distribution and spectrum}
Fig.~\ref{fig:SUPDensity} shows neutron spatial distribution at the SUP, for neutrons above \SI{10}{\MeV}.
The collimation effect is clearly visible.
Fig.~\ref{fig:SUPProfile} shows neutron fluence profile at the SUP.
In Fig.~\ref{fig:SUPProfile} simulated data were folded in energy with the neutron fission cross section for \U238~\cite{tbd}, replicating measurement with a \U238\ TFBC~\cite{Prokofiev2009}.
Within the beam umbra the predicted neutron fluence rate above \SI{10}{\MeV} at a proton current of \SI{200}{\nA} is \SI{0}{\neutron\per\cm\squared\per\second}, compared to \SI{0}{\neutron\per\cm\squared\per\second} from earlier calculations~\cite{Platt13} and \SI{0}{\neutron\per\cm\squared\per\second} from the parameterisation~\cite{Prokofiev2009}.
The parameterisation therefore exceeds the current calculations by tbd\%.\todo{Quantify integral yield}

\begin{figure}[t]
    \includegraphics[width=3in]{SUP10ColSpatialDistribution10MeVRADECS.png}
    \caption{
        Spatial distribution of neutrons above \SI{10}{\MeV} at the SUP.
        Concentric white circles indicate nominal umbra and penumbra limits.
    }
    \label{fig:SUPDensity}
\end{figure}

\begin{figure}[t]
    \includegraphics[width=3in]{SUP10beamproFoldingRADECS.png}
    \caption{
        Neutron profile at the SUP, normalised to fluence rate on axis.
        Vertical lines indicate nominal umbra and penumbra limits.
    }
    \label{fig:SUPProfile}
\end{figure}
